---
title: "Interaction with time"
author: "A. Abdias Baldiviezo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(dygraphs)
library(tidyquant)
library(timetk)
library(lubridate)
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
tickers_today <- c("CXW", "F", "GM", "JCP", "KR", "WDC", "NKE","T", "WDAY", "WFC", "WMT")
multpl_stocks <- tq_get(tickers_today, from = "2015-06-01", to = "2020-06-01", get = "stock.prices")
```

## Background

Your data science income has ballooned, and you need to find somewhere to invest $25,000 that you have saved over the last year. You have a savvy investment friend that is kind enough to tell you ten stocks he has been watching during the last year. You will need to visualize the last five years performance to help in the conversation with your friend.

Your friend is going to give you his tickers at the beginning of your half-hour meeting with him (he is a busy friend). You will need to build an .Rmd file that will build a suite of visualizations rapidly for your conversation. You will need a series of stock performance graphics using library(dygraphs) and library(ggplot2). In real life, you might use one or the other, but this task is built to help you move from ts and xts to tidy objects for visualization.


## Data Wrangling

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
filtered_multpl_stocks <- multpl_stocks %>% select(symbol, date, adjusted) %>% pivot_wider(names_from = symbol, values_from = adjusted)
```

## Data Visualization

### 5 year Timeline

Adjusted value shown for each stock
<br>
<div id="labels"></div>

```{r, fig.width= 9, fig.height= 4}
# Use this R-Chunk to plot & visualize your data!

stocks <- cbind(filtered_multpl_stocks$CXW, filtered_multpl_stocks$F, filtered_multpl_stocks$GM, filtered_multpl_stocks$JCP, filtered_multpl_stocks$KR, filtered_multpl_stocks$WDC, filtered_multpl_stocks$NKE, filtered_multpl_stocks$T, filtered_multpl_stocks$WDAY, filtered_multpl_stocks$WFC, filtered_multpl_stocks$WMT, as.xts(filtered_multpl_stocks$date))

dygraph(stocks, main = "Multiple stocks from 06/2015-06/2020", ylab = "adjusted stock price") %>% dySeries("filtered_multpl_stocks.CXW", label = "CXW") %>%
  dySeries("filtered_multpl_stocks.F", label = "F") %>%
  dySeries("filtered_multpl_stocks.GM", label = "GM") %>%
  dySeries("filtered_multpl_stocks.JCP", label = "JCP") %>%
  dySeries("filtered_multpl_stocks.KR", label = "KR") %>%
  dySeries("filtered_multpl_stocks.WDC", label = "WDC") %>%
  dySeries("filtered_multpl_stocks.NKE", label = "NKE") %>%
  dySeries("filtered_multpl_stocks.T", label = "T") %>%
  dySeries("filtered_multpl_stocks.WDAY", label = "WDAY") %>%
  dySeries("filtered_multpl_stocks.WFC", label = "WFC") %>%
  dySeries("filtered_multpl_stocks.WMT", label = "WMT") %>%
  dyLegend(width = 100, labelsDiv = "labels")
```

### Volume

Trading volume is a measure of how much of a given financial asset has traded in a period of time. For stocks, volume is measured in the number of shares traded and, for futures and options, it is based on how many contracts have changed hands.
source: https://www.investopedia.com/articles/technical/02/010702.asp

```{r plot_data, fig.height= 25}
ggplot(multpl_stocks, aes(date, volume/1000, fill = symbol)) + geom_histogram(stat = "identity") + facet_wrap(~symbol, scales = "free_y", nrow = 6) + labs(y="volume in thousands")
```

```{r plot_data2, fig.height= 25}
multpl_stocks %>% plot_time_series(date, volume/1000, .color_var = symbol, .interactive = FALSE) + facet_wrap(~symbol, nrow = 11, scales = "free_y") + labs(y="volume in thousands")
```

## Conclusions

After reviewing the data and visualizations it is observable that WFC, GM and F are the stocks that are gradually growing and seem promissing. In specific WFC has a faster growth rate for 2020.
