---
title: "Interacting with space"
author: "A. Abdias Baldiviezo"
date: "July 11, 2020"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(sf)
library(ggrepel)
library(downloader)
library(leaflet)
```

```{r load_data}
# first shp for the whole map of the us
tmp <- tempfile(fileext = ".zip")
download(url = "https://byuistats.github.io/M335/data/shp.zip", destfile = tmp, mode="wb")
unzip(as.character(tmp), list = TRUE)
unzip(as.character(tmp))
states <- st_read("County-AK-HI-Moved-USA-Map.shp")
unlink(tmp)
rm(tmp)

# water map
tmp2 <- tempfile(fileext = ".zip")
download(url = "https://research.idwr.idaho.gov/gis/Spatial/Hydrography/streams_lakes/c_250k/hyd250.zip", destfile = tmp2, mode="wb")
unzip(as.character(tmp2), list = TRUE)
unzip(as.character(tmp2))
water <- st_read("hyd250.shp")
unlink(tmp2)
rm(tmp2)

# well
tmp3 <- tempfile(fileext = ".zip")
download(url = "https://opendata.arcgis.com/datasets/1abb666937894ae4a87afe9655d2ff83_1.zip", destfile = tmp3, mode="wb")
unzip(as.character(tmp3), list = TRUE)
unzip(as.character(tmp3))
well <- st_read("Wells.shp")
unlink(tmp3)
rm(tmp3)

# water
tmp4 <- tempfile(fileext = ".zip")
download(url = "https://data-idwr.opendata.arcgis.com/datasets/e163d7da3b84410ab94700a92b7735ce_0.zip", destfile = tmp4, mode="wb")
unzip(as.character(tmp4), list = TRUE)
unzip(as.character(tmp4))
dam <- st_read("Dams_of_Idaho.shp")
unlink(tmp4)
rm(tmp4)
```

## Background

We have been asked by the state of Idaho to visualize permitted well locations with a production of more than 5000 gallons and the spatial relationship of the large wells to the locations of large dams (surface area larger than 50 acres). They have provided a a shp file (https://byuistats.github.io/M335/data/shp.zip) for us and given us a web address for the well (https://data-idwr.opendata.arcgis.com/datasets/1abb666937894ae4a87afe9655d2ff83_1.zip), dam (https://data-idwr.opendata.arcgis.com/datasets/e163d7da3b84410ab94700a92b7735ce_0.zip), and water (https://research.idwr.idaho.gov/gis/Spatial/Hydrography/streams_lakes/c_250k/hyd250.zip) locations. They would like to have the Snake River and Henrys Fork rivers plotted.

Their US map is projected to show the entire US. You will need to subset the data to Idaho and then reproject the map, so Idaho is not tilted.
## Data Wrangling

```{r tidy_data}
#narrow just to Idaho
idaho <- filter(states, StateName == "Idaho")
#flip the tilted initial layout
idaho <- st_transform(idaho, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
# narrow water to only 2 rivers
rivers <- filter(water, FEAT_NAME == "Snake River" | FEAT_NAME == "Henrys Fork")
#re-project rivers
rivers <- st_transform(rivers, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

#filter large wells
summary(well$Production)
well <- filter(well, Production >= 5000)

#filter large dams
summary(dam$SurfaceAre)
dam <- filter(dam, SurfaceAre >= 50)
```

## Data Visualization

```{r plot_data}
greenLeafIcon <- makeIcon(
  iconUrl = "http://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)
m <- leaflet(idaho$geometry) %>% addTiles() %>% addMarkers(data = well$geometry, popup = well$WellUse) %>% addMarkers(data = dam$geometry, icon = greenLeafIcon, popup = dam$DamName)
m
```

## Conclusions

Here we can see in a very interesting way the dams(green) and wells(blue) in Idaho which are filtered to show dams of over 50 Acres in size and wells with production of 5000 gallons a year. You can see the name of the dam on click, as well as the ID number of the well.