---
title: "Visualization Critique and Slack Set-up"
author: "A. Abdias Baldiviezo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)

```

```{r load_data}
# Use this R-Chunk to import all your datasets!

```

## Background

[ ] Find 4-5 examples of data-driven answers and write a one-paragraph review of each.

    [ ] List 2-3 items that are unique/good
    [ ] Identify 1 issue with the each example



## Conclusions

#Example 1
After doing some reading I found this: "Google created the People Analytics Department to help the company make HR decisions using data, including deciding if managers make a difference in their teams’ performance. The department used performance reviews and employee surveys to answer this question. Initially, it appeared managers were perceived as having a positive impact. However, a closer look at the data revealed teams with better managers performed best, are happier and work at Google longer."
-I think that the way in how they colleccted teh data was efficient, who else is better reviewing their boss than the employees?
Issue: Maybe some employees may have some bias in certain ocassions
<https://programs.online.utica.edu/articles/data-driven-decisions>

#Example 2
Another Tech Company using data to ensure good decisions is Facebook, Often when they test their features they updated the app for a feature for a couple of hours or even days and based on that they decide whether to keep the feature or not, for example: "acebook looked at how many people started to use a feature through which they could ask a friend to take down a photo of them, but then abandoned that request when they realized they had to write a message to the friend explaining why. Facebook analysts found that if they auto-populated a sample message, the number of users who completed the request rose to 60 percent from 20 percent. That data drove the decision to make the auto-populated message an official part of the tool."
Issue: Updating the app multiple tiems can be burdensome for the users
<https://www.smartsheet.com/data-driven-decision-making-management#who-uses-data-driven-decision-making-examples-and-case-studies>

#Example 3
The same happens with Netflix, many of their algorithms take into consideration the data that the users produce, and it is very evident when you use this platform. After watching a movie you usually get suggestions of similar moveis or movies of the same genre.
"The company used data analytics to build a highly detailed picture of consumers’ tastes in videos. Then, they invested heavily in developing content that ticked off the viewer preference boxes."
Issue: some users might like to explore different styles and genres.
<https://www.smartsheet.com/data-driven-decision-making-management#who-uses-data-driven-decision-making-examples-and-case-studies>

#Example 4
An example/question that I found in is the following: "According to our own research of 1,000 companies, the average response time to respond to customer service requests is 12 hours and 10 minutes. That’s too long! So, how can you reduce response times?"
This question uses the data in order to improve efficiency in customer support for any company.
Issue: Teh costs of reducing time can be higher.
<https://www.superoffice.com/blog/response-times/>
