---
title: "The US Grocer (KR)"
author: "A. Abdias Baldiviezo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(dygraphs)
library(tidyquant)
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
kr_five_years <- tq_get("KR", get = "stock.prices", from = "2015-06-26", to = "2020-06-26") 

kr_two_years <- tq_get("KR", get = "stock.prices", from = "2018-04-15", to = "2020-04-15") 
```

## Background

Interactive visualizations can often be powerful for data investigations. They let your team get at questions in real time. Use the Kroger (KR) stock price to build an interactive visualization of their performance over the last five years. Take some time to build a couple of different visualizations that show Krogers performance. See the task items below for a description.

## Dygraphs Notes

-You can use dygraphs at the R console, within R Markdown documents, and within Shiny applications. 
-dygraph(data) to create
-dygraph(data) %>% dyRangeSelector() to zoom range
-dySeries("", label = "") for time series 
-dyOptions(stackedGraph = TRUE) stack graphs
-dyRangeSelector(height = 20) adjust height

## Data Wrangling

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
kr_five_years <- kr_five_years %>% tq_transmute(adjusted, periodReturn, period = "daily", col_rename = "returns")  %>% mutate(cr = cumprod(1 + returns), cumulative_returns = cr - 1 )

kr_two_years <- kr_two_years %>% tq_transmute(adjusted, periodReturn, period = "daily", col_rename = "returns")  %>% mutate(cr = cumprod(1 + returns), cumulative_returns = cr - 1 )
kr_final <- kr_two_years %>% select(date, returns, cr, cumulative_returns ) %>% mutate(cr = cr * 10000, cumulative_returns = cumulative_returns * 10000)
```

## Data Visualization
### 5 year data
Here we see the return based on 1 unit of investment and how it oscilates through the 5 year range.
```{r, fig.width= 9, fig.height= 4}
# 5 year data
dygraph(cbind(kr_five_years$cumulative_returns, as.xts(kr_five_years$date)), main = "Cumulative Return of KR stock from 2015-2020", ylab = "return amount") %>%  dySeries("kr_five_years.cumulative_returns", label = "return") %>% dyLegend(show = "always", hideOnMouseOut = FALSE)
```

### 2 year investment
With an investment of 10,000$ we can see at every point in time (daily), how much of our investment we are getting back.
In red the lowest values reached are highlighted, and in green the highest values.

```{r, fig.width= 9, fig.height= 4}
# 5 year data
dygraph(cbind(kr_final$cumulative_returns, as.xts(kr_final$date)), main = "Cumulative Return of $10,000 in KR stock 04/2015 - 04/2020", ylab = "return amount") %>%  dySeries("kr_final.cumulative_returns", label = "return") %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>% dyRangeSelector() %>%  dyShading(from = "2019-07-01", to = "2019-10-01", color = "#FFE6E6") %>%
  dyShading(from = "2020-01-01", to = "2020-04-01", color = "#CCEBD6")
```