---
title: "Strings and Regular Expression"
author: "A. Abdias Baldiviezo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: show
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
#libraries!
library(tidyverse)
library(stringr)
library(readr)
```

```{r load_data}
# datasets
rand_letters <- read_lines("https://byuistats.github.io/M335/data/randomletters.txt")
rand_numbers_and_letters <- read_lines("https://byuistats.github.io/M335/data/randomletters_wnumbers.txt")
```

## Background

Using global regular expression print (grep) and regular expressions (regex) to find character string patterns is a valuable tool in data analysis and is available with all operating systems and many different programming languages. It is a powerful tool once it is understood. The recently developed library(stringr) package makes these tools much easier to use. The three tasks below can be completed in many different ways. As a challenge, my code to complete this entire task less than 10 lines.

## Data Wrangling

```{r tidy_data}
# With the randomletters.txt file, pull out every 1700 letter (e.g. 1, 1700, 3400, …) and find the quote that is hidden - the quote ends with a period
letters_length <- str_length(rand_letters)
mysequence <- c(seq(0,letters_length, by = 1700))
mysequence[1] <- 1
result <- c()
for (i in 1:length(mysequence)) {
  char <- str_sub(rand_letters, mysequence[i], mysequence[i])
  if (char == ".") {
    result[i] <- char
    break
  }
  result[i] <- char
}
result <- str_c(result, collapse = "")
#secret phrase
result
# With the randomletters_wnumbers.txt file, find all the numbers hidden and convert those numbers to letters using the letters order in the alphabet to decipher the message
onlynumbers <- str_match_all(rand_numbers_and_letters, "([1-9][0-9]|[0-9])")
onlynumbers
alphabet <- c(letters[1:26])
trans <- c()
for (i in 1:length(onlynumbers[[1]])) {
  trans[i] <- alphabet[as.numeric(onlynumbers[[1]][i])]
}
trans <- str_c(trans, collapse = "")
#phrase
trans
#With the randomletters.txt file, remove all the spaces and periods from the string then find the longest sequence of vowels
strip <- str_replace_all(rand_letters, "([.]|[ ])", "")
#check
str_locate_all(strip, "([.]|[ ])")
maxseq <- str_locate_all(strip, "([a|e|i|o|u]+)")
maxseq <- max(maxseq[[1]])
#maximun sequence of vowels
maxseq
```
