---
title: "World Data Investigations - Part 2"
author: "A. Abdias Baldiviezo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(ourworldindata)
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
financing_healthcare <- financing_healthcare
#View(financing_healthcare)
```

## Background


    [ ] Take notes on your reading of the specified ‘R for Data Science’ chapter in the README.md or in a ‘.R’ script in the class task folder
    [ ] Explore the world data on financing_healthcare and create an interesting graphic that highlights this dataset in relation to child mortality.
        [ ] Use devtools::install_github("drsimonj/ourworldindata") to get a larger dataset on financing_healthcare
        [ ] Create a graphic that you could use in a final presentation to summarize world child mortality (child_mort column)
    [ ] Push your .Rmd, .md, and .html to your GitHub repo

## R for Data Science - Chapter 7

-A variable is a quantity, quality, or property that you can measure.

-A value is the state of a variable when you measure it. The value of a variable may change from measurement to measurement.

-An observation is a set of measurements made under similar conditions (you usually make all of the measurements in an observation at the same time and on the same object). An observation will contain several values, each associated with a different variable. I’ll sometimes refer to an observation as a data point.

-Tabular data is a set of values, each associated with a variable and an observation. Tabular data is tidy if each value is placed in its own “cell”, each variable in its own column, and each observation in its own row.

-Variation is the tendency of the values of a variable to change from measurement to measurement

-A variable is categorical if it can only take one of a small set of values
-A variable is continuous if it can take any of an infinite set of ordered values

-If you wish to overlay multiple histograms in the same plot, I recommend using geom_freqpoly() instead of geom_histogram(). geom_freqpoly() performs the same calculation as geom_histogram()
-To make it easy to see the unusual values, we need to zoom to small values of the y-axis with coord_cartesian()

-Drop the entire row with the strange values diamonds2 <- diamonds %>% filter(between(y, 3, 20))
-replacing the unusual values with missing values. The easiest way to do this is to use mutate() to replace the variable with a modified copy. You can use the ifelse() function to replace unusual values with NA diamonds2 <- diamonds %>% mutate(y = ifelse(y < 3 | y > 20, NA, y))

-Covariation is the tendency for the values of two or more variables to vary together in a related way

-If you have long variable names, geom_boxplot() will work better if you flip it 90°. You can do that with coord_flip()

## Data Wrangling

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!
summary(financing_healthcare)
colSums(is.na(financing_healthcare))
insurance <- filter(financing_healthcare, health_insurance != is.na(health_insurance) | no_health_insurance != is.na(no_health_insurance))
#View(insurance)
#colSums(is.na(financing_healthcare_no_na))
```

## Data Visualization

```{r plot_data}
# General Child Mortality
ggplot(financing_healthcare, aes(x=child_mort, color= continent, fill = continent)) + geom_histogram(binwidth = 20) + geom_vline(aes(xintercept = mean(subset(child_mort, continent == "Africa"), na.rm = TRUE), color="Africa")) +
geom_vline(aes(xintercept = mean(subset(child_mort, continent == "Americas"), na.rm = TRUE), color="Americas")) +
geom_vline(aes(xintercept = mean(subset(child_mort, continent == "Asia"), na.rm = TRUE), color="Asia")) +
geom_vline(aes(xintercept = mean(subset(child_mort, continent == "Europe"), na.rm = TRUE), color="Europe")) +
geom_vline(aes(xintercept = mean(subset(child_mort, continent == "Oceania"), na.rm = TRUE), color="Oceania")) +
scale_x_continuous(breaks = seq(0,600, by = 50)) + labs(title = "Child Mortality in each continent with the mean as a vertical line")
#no health insurance USA
ggplot(insurance, aes(x=year, y = no_health_insurance)) + geom_line() + scale_x_continuous(breaks = seq(1950,2020, by = 10)) + labs(title = "Percentage of People without healh insurance in the USA")
# health insurance coverage Europe
ggplot(insurance, aes(x=year, y = health_insurance, color = country)) + geom_line() + xlim(1900,1975) + labs(title = "Percentage of People with Health Insurance In Europe")
```

## Conclusions

We can observe that the mean (average) for Africa (~350) and Asia (~320) are the highest.
Also the data is distributed most frequently in the ~400 for most of the continents. Meaning that most of the countries in that continent ahve around that score.
