---
title: "Give Your Visualization Wings to Fly"
author: "A. Abdias Baldiviezo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    toc: true
    toc_float: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package
library(tidyverse)
library(nycflights13)
```

```{r load_data}
# Use this R-Chunk to import all your datasets!
flights <- flights
airlines <- airlines
View(airlines)
```

## Background

You just started your internship at a big firm in New York, and your manager gave you an extensive file of flights that departed JFK, LGA, or EWR in 2013. From this data (nycflights13::flights), your manager wants you to answer the following questions:

    1.If I am leaving before noon, which two airlines do you recommend at each airport (JFK, LGA, EWR) that will       have the lowest delay time at the 75th percentile?
    2.Which origin airport is best to minimize my chances of a late arrival when I am using Delta Airlines?
    3.Which destination airport is the worst (you decide on the metric for worst) airport for arrival time?
    

Tasks

    [ ] Address at least two of the three questions in the background description (if you have time try to tackle all three)
    [ ] Make sure to include one or more visualization that shows the complexity of the data. (i.e. avoid bar graphs of summaries only)
    [ ] Create one .rmd file that has your report
        [ ] Have a section for each question
        [ ] Make sure your code is in the report but defaults to hidden
        [ ] Write an introduction section that describes your results
        [ ] Make a plot of the data to show the answer to the specific question
    [ ] Push your .Rmd, .md, and .html to your GitHub repo
    [ ] Be prepared to discuss your analysis in the upcoming class
    [ ] Complete the recommended reading on posting issues.
    [ ] Find two other student’s compiled files in their repository and provide feedback using the issues feature in GitHub (If they already have three issues find a different student to critique). Just leave the issue at the top level of the repository.
    [ ] Address 1-2 of the issues posted on your project and push the updates to GitHub



## Data Wrangling

```{r tidy_data}
# question 1
# get all flights before noon by airport
flights_before_noon <- filter(flights, hour < 12)
summary(flights_before_noon)
jfk_before_noon <- filter(flights_before_noon, origin == "JFK")
ewr_before_noon <- filter(flights_before_noon, origin == "EWR")
lga_before_noon <- filter(flights_before_noon, origin == "LGA")
#question 2
jfk_origin_delta <- filter(flights, origin == "JFK" & carrier == "DL")
ewr_origin_delta <- filter(flights, origin == "EWR" & carrier == "DL")
lga_origin_delta <- filter(flights, origin == "LGA" & carrier == "DL")


```

## Data Visualization

```{r plot_data}
#JFK
ggplot(data = jfk_before_noon, aes(x = dep_delay, color = carrier)) + geom_histogram(binwidth = 1) + xlim(-10,10) + labs(title = "JFK before noon flights by carrier with line at the 75% percentile") + geom_vline(aes(xintercept = quantile(dep_delay,0.75, na.rm = TRUE)))
#EWR
ggplot(data = ewr_before_noon, aes(x = dep_delay, color = carrier)) + geom_histogram(binwidth = 1) + xlim(-10,10) + labs(title = "EWR before noon flights by carrier with line at the 75% percentile") + geom_vline(aes(xintercept = quantile(dep_delay,0.75, na.rm = TRUE)))
#LGA
ggplot(data = lga_before_noon, aes(x = dep_delay, color = carrier)) + geom_histogram(binwidth = 1) + xlim(-10,10) + labs(title = "LGA before noon flights by carrier with line at the 75% percentile") + geom_vline(aes(xintercept = quantile(dep_delay,0.75, na.rm = TRUE)))
```

## Question 1

If I am leaving before noon, which two airlines do you recommend at each airport (JFK, LGA, EWR) that will  have the lowest delay time at the 75th percentile?

Response: as we can see in the graphs which is already filtered for the data of all flights before noon, the  two most recommendable airlines are:
<br>
-JFK -> VX (Virgin America) and US (US Airways Inc.)<br>
-EWR -> VX (Virgin America) and WN (Southwest Airlines Co.)<br>
-LGA -> YV (Mesa Airlines Inc.) and WN (Southwest Airlines Co.)
  
As marked by the black vertical line representing the 75% percentile.

```{r plot_data2}
# Use this R-Chunk to plot & visualize your data!
#JFK
ggplot(data = jfk_origin_delta, aes(x = arr_delay)) + geom_histogram(binwidth = 50) + scale_y_continuous(breaks = seq(0, 20000, by = 1000)) + labs(title = "JFK delay in arrivals flying with Delta Airlines")
#EWR
ggplot(data = ewr_origin_delta, aes(x = arr_delay)) + geom_histogram(binwidth = 50) + scale_y_continuous(breaks = seq(0, 20000, by = 1000)) + labs(title = "EWR delay in arrivals flying with Delta Airlines")
#LGA
ggplot(data = lga_origin_delta, aes(x = arr_delay)) + geom_histogram(binwidth = 50) + scale_y_continuous(breaks = seq(0, 20000, by = 1000)) + labs(title = "LGA delay in arrivals flying with Delta Airlines")


c(count(jfk_origin_delta), count(ewr_origin_delta), count(lga_origin_delta))
```

## Question 2
Which origin airport is best to minimize my chances of a late arrival when I am using Delta Airlines?

Response:
There is this number of flights in time in each origin airport flying with delta, compared with the number of flights:<br>
-JFK: approx 13400 of 20559<br>
-EWR: approx 3800 of 4295<br>
-LGA: approx 16700 of 22804<br>
dividing and getting the accuracies we can conclude that:<br>
-JFK: approx 65%<br>
-EWR: approx 88%<br>
-LGA: approx 73%<br>
The best airport to avoid arrival delays is EWR


